The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2017)

### 2018-11-30 outputs/docs/events/paper.Rmd
Update paper with Matt's commetns

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2017)"
thanks: "Thank you to Ben Readshaw, Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU School of Politics and International Relations, the ANU Research School of Economics, the Australian Parliamentary Library, and the U of T Political Behavior Group. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "We systematically analyse how discussion in the Australian Federal Parliament changes in response to elections and prime ministers. We first create a dataset of what was said in the Australian Federal Parliament from 1901 through to 2017 based on available public records. To reduce the dimensionality of this dataset we use a correlated topic model, and then analyse the effect of various events using a Bayesian hierarchical Dirichlet model. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and elections that do not result in a change in prime minister are rarely associated with topic changes. The effect of events is much more pronounced in the second half of our sample, and especially in the past two decades."
keywords: "text-as-data, Australian political history, unsupervised machine learning, Bayesian hierarchical Dirichlet model"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

Our work contributes to the growing literature in quantitative social science that considers text as an input to statistical methods, rather than requiring separate analysis. This literature sits across and draws from various historically-separate disciplines including applied statistics, economics, and political science. We contribute to this literature in terms of both data and methods. From a data perspective we bring to bear an essentially-complete record of what was said in the Australian Federal Parliament on a daily basis, and we make our dataset available to other researchers via the R package \texttt{hansardr}. From a methods perspective, our analysis model has several advantages over existing methods. These include: allowing the events to have more-complicated auto-correlated functional forms; implementing pooling across groups of similar documents; and identifying outlying topic distributions without the need to pre-specify the event of interest. As the digitisation of historical sources continues and computational power becomes cheaper, we expect the popularlity of quantitative approaches to text to further increase and the data and methods in our paper can be improved on in a variety of ways.

@@

The percentage of stop-words in each record is reasonably consistent over time. This suggests that there is no significant difference in the quality of the parsing over time, although manual inspection does suggest there is some improvement in quality over time. Details of this check are provided in Appendix \@ref(stopwordsgraph). We use Hansard records on a daily basis in this paper. We pre-process our text before applying a topic model. The specific steps that we take are to: remove numbers and punctuation; change the words to lower case; and concatenate multi-word names titles and phrases, such as new zealand to new_zealand. Then the sentences are de-constructed and each word considered individually. The resulting dataset used for analysis contains counts of words by day for each of the 14,551 days between 1901 and 2017.

@@

There are several outputs of interest from this modelling stage. For example, the model provides estimates of topic prevalence by each sitting period. This nicely illustrates how the topics change over time, as the daily estimates tend to be quite variable, but using periods defined by prime ministers or elections tend not to provide enough variability. For instance, examining Topics 26, 28, 30 and 66, which have to do with war and conflict illustrates Australia's involvement in World War I, World War II, the Korean War, the Vietnam War, the First Gulf War, the War in Afghanistan and the Second Gulf War (Figure \@ref(fig:wartopicsgraph)).

@@

# Discussion
Of the 36 prime ministers over this period, we find that 20 of them are significantly different to the prime minister that preceded them. However, four of these results -- the significance of the Page Premiership, the First Menzies Premiership, the Chifley Premiership and the Abbott Premiership -- are likely due to the short length of either the prime minister or the predecessor and should be ignored, leaving only 16. The earliest of these are the First Fisher Premiership, the Cook Premiership and the Third Fisher Premiership, which were different to the Second Deakin Premiership, the Second Fisher Premiership, and the Cook Premiership, respectively. To a certain extent this is likely due to changes due to World War I.

@@

In this paper we think of events as affecting discussion in parliament. Given the events that we consider and the broad topics of discussion we consider this the most appropriate direction for this paper. Other papers could consider an alternative specification, especially if they were to focus on particular aspects.

@@

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2017)"
thanks: "Thank you to Ben Readshaw, Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU School of Politics and International Relations, the ANU Research School of Economics, the Australian Parliamentary Library, and the U of T Political Behavior Group. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "We systematically analyse how discussion in the Australian Federal Parliament changes in response to elections and prime ministers. We first create a dataset of what was said in the Australian Federal Parliament from 1901 through to 2017 based on available public records. To reduce the dimensionality of this dataset we use a correlated topic model, and then analyse the effect of various events using a Bayesian hierarchical Dirichlet model. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; elections that do not result in a change in prime minister are rarely associated with topic changes; and that these effects have become more pronounced since the 1980s."
keywords: "text-as-data, Australian political history, unsupervised machine learning, Bayesian hierarchical Dirichlet model"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

Our work contributes to the growing literature in quantitative social science that considers text as an input to statistical methods, rather than requiring separate analysis. This literature sits across and draws from various historically-separate disciplines including applied statistics, economics, and political science. We contribute to this literature in terms of both data and methods, as well as our key findings. From a data perspective we bring to bear an essentially-complete record of what was said in the Australian Federal Parliament on a daily basis, and we make our dataset available to other researchers via the R package \texttt{hansardr}. From a methods perspective, our analysis model has several advantages over existing methods. These include: allowing the events to have more-complicated auto-correlated functional forms; implementing pooling across groups of similar documents; and identifying outlying topic distributions without the need to pre-specify the event of interest. Those contributions allow us to to find that the effects in Australia of elections and prime ministers appears to have become more pronounced since the 1980s. As the digitisation of historical sources continues and computational power becomes cheaper, we expect the popularlity of quantitative approaches to text to further increase and the data and methods in our paper can be improved on in a variety of ways.

@@

The percentage of stop-words in each record is reasonably consistent over time. This suggests that there is no significant difference in the quality of the parsing over time, although manual inspection does suggest there is some improvement in quality over time. Details of this check are provided in Appendix \@ref(stopwordsgraph). We use Hansard records on a daily basis in this paper. We pre-process our text before applying a topic model. The specific steps that we take are to: remove numbers and punctuation; change the words to lower case; and concatenate multi-word names titles and phrases, such as new zealand to new_zealand. Then the sentences are de-constructed and each word considered individually. The resulting dataset used for analysis contains counts of words by day for each of the 14,551 days between 1901 and 2017. We do not stem the words because **[Add that reference that says that it affects the results.]**

@@

There are several outputs of interest from this modelling stage. For example, the model provides estimates of topic prevalence by each sitting period. This nicely illustrates how the topics change over time, as the daily estimates tend to be quite variable, but using periods defined by prime ministers or elections tend not to provide enough variability. For instance, examining Topics 26, 28, 30 and 66, which have to do with war and conflict illustrates Australia's involvement in World War I, World War II, the Korean War, the Vietnam War, the First Gulf War, the War in Afghanistan and the Second Gulf War (Figure \@ref(fig:wartopicsgraph)). Appendix \@ref(topicmodelresultsdiscussion) provides more detail about the topic model results in the context of economic events.

@@

# Discussion

**[Needs more.]**

Of the 36 prime ministers over this period, we find that 20 of them are significantly different to the prime minister that preceded them. However, four of these results -- the significance of the Page Premiership, the First Menzies Premiership, the Chifley Premiership and the Abbott Premiership -- are likely due to the short length of either the prime minister or the predecessor and should be ignored, leaving only 16. The earliest of these are the First Fisher Premiership, the Cook Premiership and the Third Fisher Premiership, which were different to the Second Deakin Premiership, the Second Fisher Premiership, and the Cook Premiership, respectively. To a certain extent this is likely due to changes due to World War I.

@@

In this paper we think of events as affecting discussion in parliament. Given the events that we consider and the broad topics of discussion we consider this the most appropriate direction for this paper. Other papers could consider an alternative specification, especially if they were to focus on particular aspects. For instance, the effect of gender, the televising of question time, or the state of the economy and the budget position.

@@

# Topic model outputs - economics {#topicmodelresultsdiscussion}

**[TBD]**
